---
title: "Analysis of longitudinal data"
author: "Hem Raj Bhattarai"
date: "4.12.2019"
output: html_document
---

# Chapter6. Analysis of longitudinal data
## Graphical Display and summary measure approach on data "RATS"
1. This chapter is all about analyses of response data/variable that is collected over different occasions over certain period of time from the same subject/individual which is in g
eneral called repeated measurement. In repeated measurement the data is recorded in longitudinal form.
2. Since the response variable is collected on same individual over time, it is likely that these response variable are dependent or correlated to each other. Additionally, the effect of explanatory variables on response variable must be consider while analyses in such data.
3. Therefore, in longitutinal data to account the dependency of response variable and effect of explanaotry varibles on response variable, we must use a suitable model. However, before that data exploration and knowing the patterns in the data is very important which is the main task of this section.
4. In this chapter we will work on ***Graphical display*** of individual response and initial assessment of data via ***summary measure method***. Under the latter section, we will focus on **choosing summary measures** where we will incorporate the **outcome on pre-treatment values** and also deal with the **missing values**. 

## Read the data and load the necessary packages
```{r}
RATSL <- read.table("~/IODS-project/data/RATSL.txt")
RATSL$ID <- factor(RATSL$ID )       #Converting categorical variable (ID) to a factor variable.
RATSL$Group <- factor(RATSL$Group)  #Converting categorical variable (Group) to a factor variable.
dim(RATSL)
str(RATSL)
head(RATSL)
library(dplyr)
library(tidyr)
library(ggplot2)
```


## Graphical Display-original data
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(RATSL, aes(x = Time, y = Weight, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(RATSL$Weight), max(RATSL$Weight)))
```
### Interpretation of the data

## Graphical Display- standarized data
```{r echo=FALSE, message=FALSE, warning=FALSE}
RATSL <- RATSL %>% group_by(Time) %>% mutate(stdwt = Weight) %>% ungroup() ## Standardise the variable Weight
glimpse(RATSL)

ggplot(RATSL, aes(x = Time, y = stdwt, linetype = ID)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ Group, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(name= "standarized weight" )

```

## Summary measure method
```{r}
n <- RATSL$Time %>% unique() %>% length()
RATSS <- RATSL %>%
group_by(Group, Time) %>%
summarise( mean=mean(Weight), se=sd(Weight)/sqrt(n) ) %>%
ungroup()

glimpse(RATSS)
ggplot(RATSS, aes(x = Time, y = mean, linetype = Group, shape = Group)) +
  geom_line() +
  scale_linetype_manual(values = c(1,2,3)) +
  geom_point(size=3) +
  scale_shape_manual(values = c(1,2,3)) +
  #geom_errorbar(aes(ymin=mean-se, ymax=mean+se, linetype="1"), width=0.3) +
  theme(legend.position = "top") +
  geom_point(aes(color=Group))+
  scale_y_continuous(name = "mean(Weight) +/- se(Weight)")

```

## Looking for outliers-whole dataset 
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(RATSL, aes(x = factor(Time), y = Weight, fill = Group))+
  geom_boxplot(position = position_dodge(width = 0.9))+
  theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(legend.position = "top")+
  scale_x_discrete(name = "Time")

```

## Looking for an outlier-in the summarized data 
```{r echo=FALSE, message=FALSE, warning=FALSE}
RATSL64S <- RATSL %>% filter(Time > 0) %>% # creating a summary by group and ID 
  group_by(Group, ID) %>%
  summarise( mean=mean(Weight) ) %>%
  ungroup()
glimpse(RATSL64S)

ggplot(RATSL64S, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), Time 1-64")

#removing an outlier from three different groups using filter function. 
RATSL64S1<- RATSL64S %>% filter((mean > 250 & Group==1)|(mean < 500 & Group==2)|(mean > 500 & Group==3))
glimpse(RATSL64S1)

ggplot(RATSL64S1, aes(x = Group, y = mean)) +
  geom_boxplot() +
  stat_summary(fun.y = "mean", geom = "point", shape=23, size=4, fill = "white") +
  scale_y_continuous(name = "mean(Weight), Time 1-64")

```

## ANOVA on outlier excluded summary data
```{r echo=FALSE, message=FALSE, warning=FALSE}
#Running one-way anova to check the existance of difference in mean weight among the groups.
res.aov <- aov(mean ~ Group, data = RATSL64S1) 
summary(res.aov)
#Tukey post-hoc test is implied to identify the groups that are diferent to each other.
TukeyHSD(res.aov)
#Diagnositc plot for anova. This is used to test the assumption of ANOVA, which is "ANOVA test assumes that, the data are normally distributed and the variance across groups are homogeneous".
plot(res.aov, 1)
#checking the homgenity of variance 
library(car) #car package contains the leveneTest function
leveneTest(mean ~ Group, data = RATSL64S1)
#ANOVA test with no assumption of equal variances
oneway.test(mean ~ Group, data = RATSL64S1)
pairwise.t.test(RATSL64S1$mean, RATSL64S1$Group,
                 p.adjust.method = "BH", pool.sd = FALSE)
plot(res.aov, 2) #Normality plot of residuals.
aov_residuals <- residuals(object = res.aov ) #Extracting the residuals
shapiro.test(x = aov_residuals ) #Run Shapiro-Wilk test on residuals
```



## Linear Mixed Effect Models (Lme) on BPRS data
1. Summary measure approach is useful first step in making inferences (a conclusion basis of evidence) about the data.
2. However, to identify the the change in repated values of response variable and to determine the explanatory variables most associated with any change there will be a need to conduct a suitable model to the data and estimating the parameters that link the explanatory variables of interest.
3. For repeated measured data, a suitable model would be a Linear mixed effect model.
## Read the data
```{r echo=FALSE, message=FALSE, warning=FALSE}
BPRSL <- read.table("~/IODS-project/data/BPRSL.txt")
BPRSL$treatment <- factor(BPRSL$treatment)       #Converting categorical variable (treatment) to a factor variable.
BPRSL$subject <- factor(BPRSL$subject)  #Converting categorical variable (subject) to a factor variable.
str(BPRSL)
```


## PLot the data
```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(BPRSL, aes(x = week, y = bprs, linetype = subject)) +
  geom_line() +
  scale_linetype_manual(values = rep(1:10, times=4)) +
  facet_grid(. ~ treatment, labeller = label_both) +
  theme(legend.position = "none") + 
  scale_y_continuous(limits = c(min(BPRSL$bprs), max(BPRSL$bprs)))
```

##Fitting LME to the BPRS data- Accounting only random intercept
```{r echo=FALSE, message=FALSE, warning=FALSE}
BPRS_reg <- lm(bprs ~ week + treatment, data = BPRSL)
summary(BPRS_reg)
library(lme4)
BPRS_ref <- lmer(bprs ~ week + treatment + (1 | subject), data = BPRSL, REML = FALSE) # Here we will fit random intercept model for the  two explanatory variables: week and treatment on bprs
summary(BPRS_ref)

```


##Fitting LME to the BPRS data- Accounting both random intercept and slope
```{r echo=FALSE, message=FALSE, warning=FALSE}
BPRS_ref1 <- lmer(bprs ~ weeks + treatment + (weeks | subject), data = BPRSL, REML = FALSE) 
summary (BPRS_ref1)
```

## Run ANOVA to find best model (BPRS_ref vs BPRS_ref1)
```{r echo=FALSE, message=FALSE, warning=FALSE}
anova(BPRS_ref, BPRS_ref1)
```

##Fitting LME to the BPRS data- Accounting both random intercept and slope with interactions between.
```{r echo=FALSE, message=FALSE, warning=FALSE}
BPRS_ref2 <- lmer(bprs ~ weeks * treatment + (weeks | subject), data = BPRSL, REML = FALSE)
summary(BPRS_ref2)
anova(BPRS_ref1, BPRS_ref2)



```




