---
title: "Chapter3"
author: "Hem Raj Bhattarai"
date: "13.11.2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Logistic Regression
## Reading the data
```{r}
setwd("~/IODS-project")
alc <- read.table("~/IODS-project/data/alc.txt") 
names(alc) 
```

### Explanation of the data
 + 
 +
 
## Development of hypothesis based on selected variables
   + There are many interesting variables in the data which could strongly link to the alcohol        consumption rate.However, I find followings varibales quite interesting and propose my personal hypothesis for each of them in reagrds with alcohol consumption. This selection was mainly based considering factors that could potential either encourage or discourage alcohol consumption. 
     + ***activities***: Having enough extra-circular activites could discourage alcohol consumption beacuse it will assist to increase the level of refreshment.
     + ***freetime***: More freetime after school could discourage alcohol consumption becuase it will assist on easing the stress that have likely accumulated during shool hours.
     + ***studytime***: More study hours demands more focus. This will likely induce the feelings of having an extra mode of refreshment, therefore could encourage alcohol consumption   
     + ***absences***:Absences may are results of stress related things, not keen on studying etc. Absences in school means that student spent their times also doing something else than just studying. Therefore, more absences in school could be a sign of spending time for drinking alcohol.
    
     
 
## Exploring the distribution of selected variables and their relationships with alcohol consumption
  + To explore the distribution of selected variables lets first install the following packages from library.
```{r}
library(ggplot2)
```
 + For graphical distribution I will use box plots of variables against the alcohol consumptions between male and female students. For that let's run following script
```{r}
g1 <- ggplot(alc, aes(x = high_use, y = activities, col = sex))
g1 + geom_boxplot() + ylab("activities")
g2 <- ggplot(alc, aes(x = high_use, y = freetime, col = sex))
g2 + geom_boxplot() + ylab("freetime")
g3 <- ggplot(alc, aes(x = high_use, y = studytime, col = sex))
g3 + geom_boxplot() + ylab("studytime")
g3 <- ggplot(alc, aes(x = high_use, y = absences, col = sex))
g3 + geom_boxplot() + ylab("absences")
```
 + Since the first box plot seems to not clear to understand, I will great the cross tabulation for clear understanding by using following script
```{r}
library(dplyr)
table1<- alc %>% group_by(activities) %>% summarise(count = n(), mean_high_use = mean(high_use), min_high_use = min(high_use), max_high_use = max(high_use))
table1
```

## Performing regression analyses on selected variables and alcohol consumption
```{r}
m <- glm(high_use ~ activities + freetime + studytime, + absences, data = alc, family = "binomial")
results <- summary(m)
results
coef(m)
OR <- coef(m) %>% exp
OR
CI<-confint(m)
CI
cbind(OR, CI)
```


## Explorig the predictive power of the model
## Cross-validation of the model
## 
