# Regression and model validation
Prior to regression and model validation, we looked deep into the data, understood it atributes and have worked with redefining name of some varibales and created a subset by seleccting variables of our interest for further analyses.  

##Reading the data
```{r}
setwd("~/IODS-project")
learning2014 <- read.table("~/IODS-project/data/learning2014.txt") 
str(learning2014) 
dim(learning2014) 
```


###Comment on structure and dimension of the data
 + The data frame contains 7 variables with 166 observations where age and points are presented as integers and attitude, deep, stra and surf are presented as numbers. 
 + Gender is presented with two factor levels, male (M) and female (F).

##Exploring the data 
```{r}
pairs(learning2014[!names(learning2014) %in% c("gender")],col=learning2014$gender)
```
 
###Interpreting the data 
  

## Create a more advanced plot matrix with ggpairs()

```{r}
library(GGally)
library(ggplot2)
ggpairs(learning2014, 
        mapping = aes(col = gender, alpha = 0.3), 
        lower = list(combo = wrap("facethist", bins = 20)))
```

#Linear Regression



*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.
